---
title: "Analysing Tooth Growth"
author: "Tom Garshol"
date: "3/12/2019"
output: html_document
---

## Overview
###This is the outline for the project :###
1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
4. Commit a conclusions and the reasonings for this.

### Load the data and perform some basic exploratory data analyses.
```{r}
library(lattice)
data(ToothGrowth)
str(ToothGrowth)
table(ToothGrowth$supp,ToothGrowth$dose)
bwplot(ToothGrowth$len ~ToothGrowth$supp | ToothGrowth$dose)
```
OJ -> Orange Juice  ---   VJ -> Vitamin C capsules
## Provide a basic summary of the data
```{r}
summary(ToothGrowth)
aggregate(ToothGrowth$len,list(ToothGrowth$dose,ToothGrowth$supp)
          ,FUN=function(x) c(x_mean = mean(x), x_sd = sd(x)))
```

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
I test the hypothesis that the two different supplements, orange juice and vitamin C, have no affect on tooth length.
```{r}
t.test(len ~ supp, data = ToothGrowth)
```
The interval contains zero, but the p-value is above 5% so it is not conclusive.

By comparing all doses with each other I can test the hypothesis that the does does not affect on the tooth length.
```{r}
d5 <- ToothGrowth[which(ToothGrowth$dose==.5),1]
d10 <- ToothGrowth[which(ToothGrowth$dose==1),1]
d20 <- ToothGrowth[which(ToothGrowth$dose==2),1]
d510_t1 <- t.test(d5, d10, paired=FALSE, var.equal=TRUE)
d510_t2 <- t.test(d5, d10, paired=FALSE, var.equal=FALSE)
d510 <- data.frame("p-value"=c(d510_t1$p.value, d510_t2$p.value),
                          "Conf-Low"=c(d510_t1$conf[1],d510_t2$conf[1]),
                          "Conf-High"=c(d510_t1$conf[2],d510_t2$conf[2]),
                           row.names=c("t1","t2"), "Dose"="[0.5..1]")
d520_t1 <- t.test(d5, d20, paired=FALSE, var.equal=TRUE)
d520_t2 <- t.test(d5, d20, paired=FALSE, var.equal=FALSE)
d520 <- data.frame("p-value"=c(d520_t1$p.value, d520_t2$p.value),
                            "Conf-Low"=c(d520_t1$conf[1],d520_t2$conf[1]),
                            "Conf-High"=c(d520_t1$conf[2],d520_t2$conf[2]), 
                            row.names=c("t1","t2"), "Dose"="[0.5..2]")
d1020_t1 <- t.test(d10, d20, paired=FALSE, var.equal=TRUE)
d1020_t2 <- t.test(d10, d20, paired=FALSE, var.equal=FALSE)
d1020 <- data.frame("p-value"=c(d1020_t1$p.value, d1020_t2$p.value),
                           "Conf-Low"=c(d1020_t1$conf[1],d1020_t2$conf[1]),
                           "Conf-High"=c(d1020_t1$conf[2],d1020_t2$conf[2]), 
                           row.names=c("t1","t2"), "Dose"="[1..2]")
doseTot <- rbind(d510,d520,d1020)
doseTot
```
Comparing all the doses reveals that their p-values are very low and the confidence intervals do not contain zero, so we can conclude that the dose does affect the tooth length.

###Conclusion
The tooth growth data confirmes that an increase in the dose of the supplement increases the tooth growth. Whether the type of supplement, vitamin C or orange juice, affects the tooth is difficult to say.

