---
title: "Statistcal Inference"
author: "Tom Garshol"
date: "3/12/2019"
output: html_document
---
##Overview##
This course project is for the Coursera class on statistical inference.

It investigates The Exponential Distribution in R and aplies The Central Limit Theorem to a thousand mean simulations.

The project illustrates through simulation the properties of the distribution of the mean of 40 exponentials by the means of:

A simulation exercise and a basic inferential data analysis.

###First, seed and variables are set.###

```{r }
set.seed(2019)

lambda <- 0.2
#n = 40 exponentials.
n <- 40
#Simu??s = 1000
runs <- 1000
```

###We do the simulations and make a plot.###

```{r }
plot(rexp(1000,lambda),
     main="Exponential distribution: lambda = 0.2, records = 1,000")
```

###Afterwards, we do a sample of MEAN compared to theoretical MEAN ###

```{r }
mns <- NULL
for (i in 1 : runs) mns <- c(mns, mean(rexp(n,lambda)))
#mean of exponential distribution is 1/lambda
1/lambda
mean(mns)
```
###We create a graph for the mean distribution.

```{r }
hist(mns, main="Mean Distribution", breaks=50)
abline(v=mean(mns),col="red")
```

###The last part is to show a sample variance compared to a theoretical variance.###

###Here is for the theoretical standard deviation###

```{r }
(1/lambda)/sqrt(n)
sd(mns)
```
###The variance ###

```{r }
((1/lambda)/sqrt(n))^2
sd(mns)^2
```
It is interesting to see the similarities in the two sets of values. They are very alike.

###Is the distribution normal ?###
```{r }
hist(mns, prob=TRUE, main="Mean Distribution", breaks=50)
lines(density(mns), lwd=3, col="red")
```

The simulation that we have created has a strong bias for normality. More samples will show this even stronger. By changing the variable "runs" to 100000 samples, the curve will be much more normalized. 